---
title: "Subsetting with bracket notation"
output: html_document
---

Here are a few pointers for using bracket notation. Personally, I find bracket notation a bit more intuitive. Keeping in mind that the left side of the comma is rows and the right side is columns, we can do something like this:

```{r eval = FALSE}
DF[some_condition, some_columns]
```

If we want multiple conditions, we can also do this:

```{r eval = FALSE}
DF[condition1 & condition2, some_columns] 
```

You can use & and | (pipe) to combine conditions. The & is a logical AND and the | (pipe) is a logical OR. You can also do negation using

*!(some_condition)*. 

You can also use numeric values to select certain rows - 

*df[ 1 , ] or df [ 1:10 , ]*. 

If you want all rows, just leave the left side of the comma empty - 

*DF[ , some_columns]*. 

If you want all columns, leave the right side of the comma emtpy - 

*DF [some_condition , ]* 

Also keep in mind that, when using bracket notation with data frames, columns are required when the first parameter is a condition so the comma is also required. Columns can be numeric or column names. If you use column names, they must be quoted. So, *c(1,3)* and *c("columnA", "columnB")* are both acceptable. 

For a single column, you can leave out *c()*. Also, *DF[ , ]\$columnA* is the same as *DF[ , "columnA"]* which is the same as *DF$columnA*. However, you can use variables inside brackets but you can't use a variable with $. 


If there is only a single value N without a comma, R returns a data frame with single brackets - 
*df[N]*
and a vector with double brackets - 
*df[[N]]* 
where N is a column number or a quoted column name.

**Adding a few examples of conditions.**

There are two things to remember when using conditions in a data.frame.

* The condition(s) go on the left side of the comma (since we're selecting rows)
* The comma is required 
```{r}
df <- data.frame(X=1:10, Y=rep(1:2, 5))
```
**Single Condition**

```{r}
df[df$Y == 2, ]
```

**Combining conditions.**

AND - using the ampersand symbol

```{r}
df[df$Y == 2 & df$X <=5, ]
```
```{r}
df[df$Y == 2 & df$X %in% c(5:10), ]
```

OR - using the pipe symbol
```{r}
df[df$Y == 2 | df$X == 5, ]
```
```{r}
df[df$Y == 2 | df$X %in% c(1,9), ]
```

**Which()**

The which() function returns a vector of indices where a given condition is true:

```{r}
which(df$Y == 2)
```
```{r}
df[which(df$Y == 2), ]
```
Sometimes you might want to save some indices for use later:

```{r}
idx <- which(df$Y == 2)

df[c(2,10), "Y"] <- 1

df[idx,]
```
Using which() is functionally the same as putting the condition directly inside the brackets. That's because when you use a logical vector for rows, the logical values get converted to row indices. So, you can use either a logical vector or a numeric vector for row indices.

```{r}
df[which(df$Y == 2), ]
```
```{r}
df[df$Y==2, ]
```

**Subsetting a Column**

You can also subset a single column using the $ sign and conditions.

```{r}
df$X[df$Y == 1]
```
```{r}
df$X[df$X < 5 | df$Y == 2]
```
